Using numerical modelling to investigate temperature and wave exposure as drivers

### 2020-01-07 thesis_chapter_1.Rmd
major house cleaning and re-organisng

---
title: "Quantifying passive kelp rafting phenomena and connectivity of nearshore communities around the coast of South Africa:Development of a 3D spectral numerical coastal-flow model"
author: 
- RM Coppin^[University of the Western Cape]
- C Rautenbach^[South African Weather Service]
- AJ Smit^[University of the Western Cape]
output:
  html_document:
    df_print: paged
    toc: yes
    theme:  simplex
  word_document:
    fig_caption: yes
    reference_docx: chapter_1_template.docx
    toc: yes
bibliography: references.bib
---

@@

surface_elevation <- readPNG("figures/surface_elevation_waves.png")
grid.raster(surface_elevation)
```

@@

Hs_Tp_elevation <- readPNG("figures/Hs_Tp_surface_elevation.png")
grid.raster(Hs_Tp_elevation)
```

>>>>

---
title: "Quantifying passive kelp rafting phenomena and connectivity of nearshore communities around the coast of South Africa:Development of a 3D spectral numerical coastal-flow model"
author: 
- RM Coppin^[University of the Western Cape]
- C Rautenbach^[South African Weather Service]
- AJ Smit^[University of the Western Cape]
output:
  word_document:
    fig_caption: yes
    reference_docx: chapter_1_template.docx
    toc: yes
  html_document:
    df_print: paged
    toc: yes
    theme:  simplex
bibliography: references.bib
---

@@

surface_elevation <- readPNG("figures/chapter_1/surface_elevation_waves.png")
grid.raster(surface_elevation)
```

@@

Hs_Tp_elevation <- readPNG("figures/chapter_1/Hs_Tp_surface_elevation.png")
grid.raster(Hs_Tp_elevation)
```

### 2020-01-07 thesis_chapter_2.Rmd
major house cleaning and re-organisng

---
title: Using numerical modelling to investigate temperature and wave exposure as drivers
  of morphological characteristics of the kelps, Ecklonia maxima and Laminaria pallida
author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
output:
  word_document:
    reference_docx: frontiers_template.docx
  pdf_document:
    fig_caption: yes
  html_document:
    df_print: paged
bibliography: references.bib
papersize: A4
tables: yes
abstract:  '**Abstract**: Macroalgal morphological variation is determined to a large extent by
  a combination of environmental factors, with wave exposure and temperature
  perhaps the main influences as they are key environmental properties to which a species becomes locally adapted.                Macroalgae have shown to exhibit different responses to different magnitudes of exposure to waves, such as reduction
  in overall size and strength increasing traits. In terms of temperature, warmer
  environments have been shown to reduce the overall size of resident and transplanted
  species. However, none of the past studies have identified specific wave and temperature
  metrics responsible for the morphological adaptation macroalgae exhibit. Past research
  has often used simple or two-dimensional models of wave exposure, which do not
  take into account important aspects of the nearshore environment such as wave breaking,
  refraction and diffraction. Furthermore, past studies have often used satellite-derived
  datasets as sources for temperature data; however, such data have been shown to have large bias when 
  applied to the nearshore environment. This study used *in situ* temperature data and wave power
  metrics calculated from a 3D-numerical model to identify specific temperature and wave metrics
  responsible for morphological adaptation of the kelp, *Ecklonia maxima* and *Laminaria pallida*. The results identify wave      exposure as the main driver of morphological adaptation while identifying specific wave metrics.
  Furthermore, the results show differences in wave metrics between species; and between
  deep and shallow populations.'

@@

Sites were chosen to represent an array of temperature and wave gradients, as indicated in figure \ref{fig:map study sites}. St. Helena Bay and Betty's Bay constituted the north-western and south-eastern boundary sites, respectively. These sites are roughly 300 km apart, lie within separate marine provinces @Smit2017, and span the majority of the south-west coast, in varying thermal and wave energy regimes. The region is dominated by kelp communities that persist in contrasting abiotic environments. The west coast region has been termed cool-temperate, which is defined as a region where mean monthly temperatures are always above 10°C and below 15°C [@Smit2013a]. East of Cape Point marks the beginning of an overlap or transition area, which is also referred to as the Benguela-Agulhas Transition Zone [@Smit2013a]. The Agulhas Marine Province is characterised by a wide temperature range of up to 7°C difference between mean monthly temperatures between summer and winter and is classified as warm-temperate [@Smit2013a].

@@

```{r map study sites, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="\\label{fig:map study sites}Map of the study area and sampling sites. The red dot indicates the position of Cape Town for sake of reference."}
map_combined
```

@@

```{r schematic, echo=FALSE, fig.cap='\\label{fig:schematic}Schematic of *Laminaria pallida* and *Ecklonia maxima* morphology [@Dyer2018-rv].', fig.height=6, fig.width=6}
library(jpeg)
library(grid)
library(gridExtra)

schematic <- readJPEG("figures/kelp_diagram.jpg")
grid.raster(schematic)
```

@@

img_wave_rose_A <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_5A.jpg")))
img_wave_rose_B <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_5B.jpg")))
img_wave_rose_C <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_5C.jpg")))

@@

wp_1 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6A.jpg")))
wp_2 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6B.jpg")))
wp_3 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6C.jpg")))
wp_4 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6D.jpg")))
wp_5 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6E.jpg")))
wp_6 <- rasterGrob(as.raster(readJPEG("figures/frontiers_fig_prep/fig_6_legend.jpg")))

@@

library(png)
library(grid)
library(jpeg)

img_fb <- readJPEG("figures/falsebay_spectrum.jpg")
grid.raster(img_fb)

@@

We provide strong support that variations in the environmental variables, particularly wave exposure variables, are driving kelp morphological characteristics around the Cape Peninsula. Morphological adaptation due to water motion may manifest in a number of ways in high wave energy environments. For instance, reduction of blade thickness, blade elongation, increase of stipe length, increase in stipe circumference and force of attachment [@FowlerWalker2006; @Wernberg2005c; @Friedland1995-vi; @Denny2002; @bekkby2014; @Denny1997] have been identified in previous studies. Although this study did not measure force of attachment, other morphological responses to temperature and wave variables were evident. Species-specific responses are evident in both wave exposure and temperature. In cool temperate environments *L. pallida* tended to show increases in certain morphological characteristics (stipe mass, stipe length and stipe diameter) while in the warm-temperate environments these were significantly lower. This was also true for deep *E. maxima* populations, which had longer, thinner stipes for kelp from the warm temperate compared to kelp from the cool temperate. Reduction in certain morphological characteristics has been attributed to temperature by @Serisawa2002-dc in the kelp *E. cava*, which was smaller and shorter in warmer sites compared to cooler sites. The reduction in size of adult may be a response to low nutrient conditions, which has been shown to reduce growth rate and overall morphology [@Simonson2015]. Warmer temperatures are associated with low nutrient concentrations [@Waldron1992] and the low frequency of upwelling conditions in False Bay (low nutrient supply) coupled with warmer temperatures may be a contributing factor. It should be noted however, that from these analyses that in general the cool temperate region is more exposed to waves than that of the warm-temperate region. Therefore, the significantly larger morphological characteristics for kelp from the cool temperate region may overlap with responses to wave exposure. The response of kelp morphological characteristics to wave exposure was evident and both species exhibit tactics based on the magnitude of wave exposure. Strength increasing traits were exhibited for lower exposure levels while a 'go-with-the-flow' tactic for moderate levels of wave exposure. This suggests that how morphological characteristics manifest themselves is dependent on the magnitude of wave exposure. When grouped by wave exposure category, *L. pallida* characteristics showed a significant increase in length for the 'extremely sheltered' category compared to the fully sheltered category, which may be a 'go-with-the-flow' tactic [@Denny1997; @Denny2002; @Hurd2000]. Kelp are able to increase flexibility by increasing stipe length which increases the extension capabilities of kelp to a passing wave [@Denny2002; @Hurd2000]. However, increasing stipe length is only beneficial under lower exposure levels as a long stipe actually increases overall drag on the plant under high wave exposure [@Denny2002; @Denny1997]. This is reflected in the *L. pallida* morphology which shows an overall reduction of morphological characteristics suggesting a size reducing tactic to cope with higher levels of wave exposure. The results suggest that deep *E. maxima* populations exhibit a different response, as well as a higher wave exposure threshold. Unlike *L. pallida*, *E. maxima* exhibits a size reducing tactic with increasing wave exposure by  decreasing stipe mass, stipe length, stipe circumference, total length and frond length. A magnitude related response has been suggested by @wernberg2010 who investigated the temporal and spatial variations of various environmental drivers of the kelp *E. radiata*. The authors identified wave exposure as the most important driver of kelp morphological characteristics and that the type of response elicited is dependent on the magnitude of wave exposure.

>>>>

---
title: Using numerical modelling to investigate temperature and wave exposure as drivers
  of morphological characteristics of the kelps, Ecklonia maxima and Laminaria pallida
author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
output:
  word_document:
    reference_docx: markdown/frontiers_template.docx
  html_document:
    df_print: paged
  pdf_document:
    fig_caption: yes
bibliography: references.bib
papersize: A4
tables: yes
abstract:  '**Abstract**: Macroalgal morphological variation is determined to a large extent by
  a combination of environmental factors, with wave exposure and temperature
  perhaps the main influences as they are key environmental properties to which a species becomes locally adapted.                Macroalgae have shown to exhibit different responses to different magnitudes of exposure to waves, such as reduction
  in overall size and strength increasing traits. In terms of temperature, warmer
  environments have been shown to reduce the overall size of resident and transplanted
  species. However, none of the past studies have identified specific wave and temperature
  metrics responsible for the morphological adaptation macroalgae exhibit. Past research
  has often used simple or two-dimensional models of wave exposure, which do not
  take into account important aspects of the nearshore environment such as wave breaking,
  refraction and diffraction. Furthermore, past studies have often used satellite-derived
  datasets as sources for temperature data; however, such data have been shown to have large bias when 
  applied to the nearshore environment. This study used *in situ* temperature data and wave power
  metrics calculated from a 3D-numerical model to identify specific temperature and wave metrics
  responsible for morphological adaptation of the kelp, *Ecklonia maxima* and *Laminaria pallida*. The results identify wave      exposure as the main driver of morphological adaptation while identifying specific wave metrics.
  Furthermore, the results show differences in wave metrics between species; and between
  deep and shallow populations.'

@@

Sites were chosen to represent an array of temperature and wave gradients, as indicated in figure \ref{fig:map}. St. Helena Bay and Betty's Bay constituted the north-western and south-eastern boundary sites, respectively. These sites are roughly 300 km apart, lie within separate marine provinces @Smit2017, and span the majority of the south-west coast, in varying thermal and wave energy regimes. The region is dominated by kelp communities that persist in contrasting abiotic environments. The west coast region has been termed cool-temperate, which is defined as a region where mean monthly temperatures are always above 10°C and below 15°C [@Smit2013a]. East of Cape Point marks the beginning of an overlap or transition area, which is also referred to as the Benguela-Agulhas Transition Zone [@Smit2013a]. The Agulhas Marine Province is characterised by a wide temperature range of up to 7°C difference between mean monthly temperatures between summer and winter and is classified as warm-temperate [@Smit2013a].

@@

```{r map, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="\\label{fig:map}Map of the study area and sampling sites. The red dot indicates the position of Cape Town for sake of reference."}
map_combined
```

@@

```{r schematic, echo=FALSE, warning=FALSE, fig.cap='\\label{fig:schematic}Schematic of *Laminaria pallida* and *Ecklonia maxima* morphology [@Dyer2018-rv].', fig.height=6, fig.width=6}
library(jpeg)
library(grid)
library(gridExtra)

schematic <- readJPEG("figures/chapter_2/fig_2.jpg")
grid.raster(schematic)
```

@@

img_wave_rose_A <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_5A.jpg")))
img_wave_rose_B <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_5B.jpg")))
img_wave_rose_C <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_5C.jpg")))

@@

wp_1 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6A.jpg")))
wp_2 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6B.jpg")))
wp_3 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6C.jpg")))
wp_4 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6D.jpg")))
wp_5 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6E.jpg")))
wp_6 <- rasterGrob(as.raster(readJPEG("figures/chapter_2/fig_6_legend.jpg")))

@@

library(grid)
library(jpeg)

img_fb <- readJPEG("figures/chapter_2/fig_7.jpg")
grid.raster(img_fb)

@@

We provide strong support that variations in the environmental variables, particularly wave exposure variables, are driving kelp morphological characteristics around the Cape Peninsula. Morphological adaptation due to water motion may manifest in a number of ways in high wave energy environments. For instance, reduction of blade thickness, blade elongation, increase of stipe length, increase in stipe circumference and force of attachment [@FowlerWalker2006; @Wernberg2005c; @Friedland1995-vi; @Denny2002; @bekkby2014; @Denny1997] have been identified in previous studies. Although this study did not measure force of attachment, other morphological responses to temperature and wave variables were evident. Species-specific responses are evident in both wave exposure and temperature. In cool temperate environments *L. pallida* tended to show increases in certain morphological characteristics (stipe mass, stipe length and stipe diameter) while in the warm-temperate environments these were significantly lower. This was also true for deep *E. maxima* populations, which had longer, thinner stipes for kelp from the warm temperate compared to kelp from the cool temperate. Reduction in certain morphological characteristics has been attributed to temperature by @Serisawa2002-dc in the kelp *E. cava*, which was smaller and shorter in warmer sites compared to cooler sites. The reduction in size of adult may be a response to low nutrient conditions, which has been shown to reduce growth rate and overall morphology [@Simonson2015]. Warmer temperatures are associated with low nutrient concentrations [@Waldron1992] and the low frequency of upwelling conditions in False Bay (low nutrient supply) coupled with warmer temperatures may be a contributing factor. It should be noted however, that from these analyses that in general the cool temperate region is more exposed to waves than that of the warm-temperate region. Therefore, the significantly larger morphological characteristics for kelp from the cool temperate region may overlap with responses to wave exposure. The response of kelp morphological characteristics to wave exposure was evident and both species exhibit tactics based on the magnitude of wave exposure. Strength increasing traits were exhibited for lower exposure levels while a 'go-with-the-flow' tactic for moderate levels of wave exposure. This suggests that how morphological characteristics manifest themselves is dependent on the magnitude of wave exposure. When grouped by wave exposure category, *L. pallida* characteristics showed a significant increase in length for the 'extremely sheltered' category compared to the fully sheltered category, which may be a 'go-with-the-flow' tactic [@Denny1997; @Denny2002; @Hurd2000]. Kelp are able to increase flexibility by increasing stipe length which increases the extension capabilities of kelp to a passing wave [@Denny2002; @Hurd2000]. However, increasing stipe length is only beneficial under lower exposure levels as a long stipe actually increases overall drag on the plant under high wave exposure [@Denny2002; @Denny1997]. This is reflected in the *L. pallida* morphology which shows an overall reduction of morphological characteristics suggesting a size reducing tactic to cope with higher levels of wave exposure. The results suggest that deep *E. maxima* populations exhibit a different response, as well as a higher wave exposure threshold. Unlike *L. pallida*, *E. maxima* exhibits a size reducing tactic with increasing wave exposure by  decreasing stipe mass, stipe length, stipe circumference, total length and frond length. A magnitude related response has been suggested by @Wernberg2010a who investigated the temporal and spatial variations of various environmental drivers of the kelp *E. radiata*. The authors identified wave exposure as the most important driver of kelp morphological characteristics and that the type of response elicited is dependent on the magnitude of wave exposure.
